---
title: "But I can't preregister my research"
subtitle: "Improving the reproducibility and transparency of ecology and conservation with adaptive preregistration for model-based research"
author:
  - id: aut1
    name:
      given: Elliot
      family: Gould
    email: elliot.gould@unimelb.edu.au
    orcid: 0000-0002-6585-538X
    attributes:
      corresponding: true
    roles:
      - conceptualization
      - data curation
      - formal analysis
      - funding acquisition
      - investigation
      - methodology
      - project administration
      - resources
      - software
      - validation
      - visualization
      - writing - original draft
      - writing - review & editing
    affiliations:
      - ref: aff1
  - id: aut2
    name:
      given: Chris
      family: Jones
    email: chris.jones@delwp.vic.gov.au
    orcid: 0000-0003-2833-0194
    attributes:
      corresponding: false
    roles:
      - conceptualization
      - data curation
      - funding acquisition
      - investigation
      - methodology
      - project administration
      - resources
      - software
      - supervision
      - validation
      - visualization
      - writing - review & editing
    affiliations:
      - ref: aff2
  - id: aut3
    name:
      given: Jian
      family: Yen
    email: jian.yen@delwp.vic.gov.au
    orcid: 0000-0001-7964-923X
    attributes:
      corresponding: false
    roles:
      - conceptualization
      - data curation
      - formal analysis
      - investigation
      - methodology
      - project administration
      - resources
      - software
      - supervision
      - validation
      - visualization
      - writing - review & editing
    affiliations:
      - ref: aff2
  - id: aut4
    name:
      given: Hannah
      family: Fraser
    email: hannahsfraser@gmail.com
    orcid: 0000-0003-2443-4463
    attributes:
      corresponding: false
    roles:
      - conceptualization
      - investigation
      - methodology
      - supervision
      - writing - review & editing
    affiliations:
      - ref: aff1
  - id: aut5
    name:
      given: Henry
      family: Wootton
    email: henry.wootton@delwp.vic.gov.au
    orcid: 0000-0001-6506-0248
    attributes:
      corresponding: false
    roles:
      - data curation
      - formal analysis
      - investigation
      - methodology
      - resources
      - software
      - visualization
      - writing - review & editing
    affiliations:
      - ref: aff2
  - id: aut6
    name:
      given: Lyndsey
      family: Vivian
    email: lyndsey.vivian@delwp.vic.gov.au
    orcid: 0000-0003-1927-723X
    attributes:
      corresponding: false
    roles:
      - data curation
      - investigation
      - methodology
      - resources
      - writing - review & editing
    affiliations:
      - ref: aff2
  - id: aut7
    name:
      given: Megan
      family: Good
    email: goodm@unimelb.edu.au
    orcid: 0000-0002-6908-1633
    attributes:
      corresponding: false
    roles:
      - investigation
      - methodology
      - writing - review & editing
    affiliations:
      - ref: aff1
  - id: aut8
    name:
      given: David
      family: Duncan
    email: david.duncan@unimelb.edu.au
    orcid: 0000-0003-4411-8214
    attributes:
      corresponding: false
    roles:
      - investigation
      - methodology
      - writing - review & editing
    affiliations:
      - ref: aff1
  - id: aut9
    name:
      given: Libby
      family: Rumpff
    email: lrumpff@unimelb.edu.au
    orcid: 0000-0001-9400-8086
    attributes:
      corresponding: false
    roles:
      - conceptualization
      - investigation
      - methodology
      - supervision
      - writing - review & editing
    affiliations:
      - ref: aff1
  - id: aut10
    name:
      given: Fiona
      family: Fidler
    email: fidlerfm@unimelb.edu.au
    orcid: 0000-0002-2700-2562
    attributes:
      corresponding: false
    roles:
      - conceptualization
      - funding acquisition
      - methodology
      - supervision
      - writing - review & editing
    affiliations:
      - ref: aff1
affiliations:
  - id: aff1
    name: MetaMelb Research, University of Melbourne
  - id: aff2
    name: Arthur Rylah Institute for Environmental Research
abstract: |
  **1.** Preregistration is an open-science practice adopted in scientific disciplines that have begun to confront the 'reproducibility crisis', which aims to improve research transparency and mitigate questionable research practices within a study. There has been little uptake of preregistration in ecology and conservation, and many modellers and researchers engaged in non-hypothesis testing research both within and outside of ecology have eschewed preregistration on the grounds that existing templates are irrelevant because they are focussed on null-hypothesis significance testing.
  **2.** We advocate that preregistration can and should be used in model-based research in ecology and conservation but that preregistration should reflect the norms and practice of the research context in which it is applied in order to adequately restrict researcher degrees of freedom that are unique to a domain-specific methodology. We took a user-centred design approach to the task of translating preregistration into model-based research contexts for ecology and conservation.
  **3.** We identified a generalised 'modelling workflow' that captures critical analysis decisions in the process of the model development cycle. The workflow informed the content and structure of a preregistration template to be used in model-based research. The iterative nature of model development was identified as a key barrier to implementing preregistration for modelling research because it conflicts with the current format and linear process of preregistration. To better align the internal logic of preregistration with the iterative and non-linear process of ecological modelling, we propose a methodology for implementing 'adaptive preregistration' leveraging git and GitHub. We tested and evaluated the preregistration template and methodology using a case-study of environmental flows modelling for decision-support in Victoria, Australia.
  **4.** This research provides a template and methodology for implementing adaptive preregistration of ecological models in either pure or applied settings, and will facilitate the wider uptake of preregistration in ecology. Although we focus on ecology and conservation in this paper, the template is extensible to alternative fields and methods, and the proposed methodology for implementing adaptive preregistration can be applied in model-based research within science more broadly. Modellers need no longer cry "but, I can't preregister my research."
keywords:
  - preregistration
  - transparency
  - reproducibility
  - modelling
  - ecology
  - conservation
  - open-science
  - research software engineering
date: last-modified
bibliography: ../../references.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
---

Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal. It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.

For detailed instructions regarding the elsevier article class, see <https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions>

# Bibliography styles

Here are two sample references: @Feynman1963118 @Dirac1953888.

By default, natbib will be used with the `authoryear` style, set in `classoption` variable in YAML. You can sets extra options with `natbiboptions` variable in YAML header. Example

```         
natbiboptions: longnamesfirst,angle,semicolon
```

There are various more specific bibliography styles available at <https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>. To use one of these, add it in the header using, for example, `biblio-style: model1-num-names`.

## Using CSL

If `cite-method` is set to `citeproc` in `elsevier_article()`, then pandoc is used for citations instead of `natbib`. In this case, the `csl` option is used to format the references. By default, this template will provide an appropriate style, but alternative `csl` files are available from <https://www.zotero.org/styles?q=elsevier>. These can be downloaded and stored locally, or the url can be used as in the example header.

# Equations

Here is an equation: $$ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
$$

Inline equations work as well: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$

# Figures and tables

@fig-meaningless is generated using an R chunk.

```{r}
#| label: fig-meaningless
#| fig-cap: A meaningless scatterplot
#| fig-width: 5
#| fig-height: 5
#| fig-align: center
#| out-width: 50%
#| echo: false
plot(runif(25), runif(25))
```

# Tables coming from R

Tables can also be generated using R chunks, as shown in @tbl-simple example.

```{r}
#| label: tbl-simple
#| tbl-cap: Caption centered above table
#| echo: true
knitr::kable(head(mtcars)[,1:4])
```

# References {.unnumbered}
