---
title: "Main Analysis: Preregistered vs Realised Models"
format:
  html:
    self-contained: true
    code-line-numbers: true
    code-copy: false
filters:
  - quarto
  - line-highlight
---

For simplicity, elements of the R model specification have been omitted, e.g. the data argument.

# Maximal Model

```{r}
#| eval: false
#| echo: true
#| filename: "maximal_model_preregistered.R"
maximal_model_preregistered <- glmmTMB::glmmTMB(
  response # cover or species richness
  ~ cover_in_prev_survey + # cover only
  wpfg * origin * flow_metrics +
  wpfg * origin * zone * period +
  grazing +
  (1 | site / transect) +
  (1 | sub_transect_height) +
  (1 | survey_year) +
  (1 | waterbody) +
  offset(npoint), # cover only
  family = poisson
  # family = nbinom2,
  # ziformula = ~ wpfg,
  # dispformula = ~ wpfg
)
```

# Vegetation Cover

## Flow Regime Models

```{r}
#| file: !expr here::here("figures/cover_ar_TMBmod_1.R")
#| eval: false
#| echo: true
#| filename: "cover_regime_preregistered.R"
#| source-line-numbers: "3,4,5,14,15,18"
```

1.  Preregistration Deviation
2.  Preregistered Flexibility

```{r}
#| file: !expr here::here("figures/cover_ar_TMBmod_1_realised.R")
#| eval: false
#| echo: true
#| filename: "cover_regime_final.R"
#| source-line-numbers: "3,4,5,9,10"
```

```{r}
#| echo: false
library(diffviewer)
path1 <- here::here("figures/cover_ar_TMBmod_1.R")
path2 <- here::here("figures/cover_ar_TMBmod_1_realised.R")

visual_diff(path1, path2)
```

## Flow Event Models

### Preregistered Model A

```{r}
#| eval: false
#| echo: true
#| filename: "cover_events_A_preregistered.R"

cover_events_A_preregistered <- glmmTMB::glmmTMB(
  hits ~ cover_in_prev_survey +
  # days_above_baseflow_std * wpfg * origin +
  # days_above_springfresh_std * wpfg * origin +
  # days_above_baseflow_std^2 +
  # days_above_springfresh_std_sq +
  zone * period +
  origin + wpfg +
  grazing +
  (1 | site / transect) +
  # (1 | site / period) +
  (1 | sub_transect_height) +
  (1 | survey_year) +
  (1 | waterbody),
  # offset(npoint),
  family = poisson,
  ziformula = ~ wpfg,
  # dispformula = ~ wpfg
)
```

### Preregistered Model B

```{r}
#| eval: false
#| echo: true
#| filename: "cover_events_B_preregistered.R"

cover_events_B_preregistered <- glmmTMB::glmmTMB(
  hits ~ cover_in_prev_survey +
    # days_above_baseflow_std * wpfg * origin +
    # days_above_springfresh_std * wpfg * origin +
    # days_above_baseflow_std^2 +
    # days_above_springfresh_std^2 +
    zone * period +
    origin + wpfg +
    grazing +
    (1 | site / transect) +
    # (1 | site / period) +
    (1 | sub_transect_height) +
    (1 | survey_year) +
    (1 | waterbody),
  # offset(npoint),
  family = poisson,
  ziformula = ~ wpfg,
  # dispformula = ~ wpfg
)
```

### Final Cover Flow Events Model

```{r}
#| eval: false
#| echo: true
#| filename: "cover_events_final.R"
cover_events_final <- glmmTMB::glmmTMB(
  hits ~ cover_in_prev_survey +
    zone * wpfg * period +
    origin +
    grazing +
    (1 | site / transect) +
    (1 | metres) +
    (1 | survey_year) +
    (1 | waterbody),
  family = poisson,
  ziformula = ~ wpfg,
  dispformula = ~ wpfg
)
```

The final cover flow events model merged the two candidate simplified models.

# Vegetation Richness

## Flow Regime Model

```{r}
#| eval: false
#| echo: true
#| filename: "richness_regime_preregistered.R"

richness_regime <- glmmTMB::glmmTMB(
  richness ~
    days_above_baseflow_std * wpfg * origin +
    days_above_springfresh_std * wpfg * origin +
    # days_above_baseflow_std^2 +
    # days_above_springfresh_std^2 +
    # zone * period +
    # zone * wpfg +
    # wpfg * period +
    # grazing + origin +
    # (1 | site / period) +
    (1 | site / transect) +
    (1 | sub_transect_height) +
    (1 | survey_year) +
    (1 | waterbody),
  family = poisson,
  # family = nbinom2,
  # ziformula = ~ wpfg,
  # dispformula = ~ wpfg
)
```

## Flow Events Model

```{r}
#| eval: false
#| echo: true
#| filename: "richness_events_preregistered.R"

richness_events <- glmmTMB::glmmTMB(
  richness ~
    # days_above_baseflow_std +
    # days_above_springfresh_std +
    # days_above_baseflow_std^2 +
    # days_above_springfresh_std^2 +
    # zone * period +
    zone * period +
    zone * wpfg +
    wpfg * period +
    grazing +
    origin +
    # (1 | site / period) +
    (1 | site / transect) +
    (1 | sub_transect_height) +
    (1 | survey_year) +
    (1 | waterbody),
  family = poisson,
  # family = nbinom2,
  # ziformula = ~ wpfg,
  # dispformula = ~ wpfg
)
```

## Final Richness Model

```{r}
#| eval: false
#| echo: true
#| filename: "richness_final.R"
#| #source-line-numbers: "4"

richness_ar_TMB_final <- glmmTMB::glmmTMB(
  richness ~
    days_above_baseflow_std * wpfg + # flow regime model, rm origin
    days_above_springfresh_std * wpfg +
    days_above_baseflow_std^2 +
    days_above_springfresh_std^2 +
    zone * wpfg * period + # maximal model, rm origin
    grazing + # flow events model
    origin +  # flow events model
    (1 | site / transect) +
    (1 | sub_transect_height) +
    (1 | survey_year) +
    (1 | waterbody),
  family = poisson
)
```

> The full dataset supported a more complex model than the pilot study, with all predictors included in a single model, and a three-way interaction between zone, sampling period, and functional group.\
> Quadratic associations with flow metrics were unable to be fitted at the functional-group level and were instead included as a single term for all functional groups.
